D.I.A.G.R.A.M.: Development of Image Analysis for Graph Recognition And Modeling

### 2025-06-29 doc/article/article.tex
Updated paper

Some works in element recognition - like arrows - are particularly domain-specific (like medical imaging \cite{med1}\cite{med2}). Other studies focus on the so-called \textit{online recognition} task, whose goal is to recognize diagrams as they are drawn (see \cite{online1}, \cite{online2}). The online recognition task quite often employs the use of relative strokes, and is mainly focused on flowchart diagrams.

@@

\section{Dataset}
TODO: Elencare i dataset usati, in che modo erano usati originariamente e come lo usiamo noi (quali dati per quali componenti) \\

>>>>

Some works in element recognition - like arrows - are particularly domain-specific (like medical imaging \cite{med1}\cite{med2}). Other studies focus on the so-called \textit{online recognition} task, whose goal is to recognize diagrams as they are drawn (see \cite{online1}, \cite{online2}, \cite{online3}). The online recognition task quite often employs the use of relative strokes, and is mainly focused on flowchart diagrams.

@@

\section{Dataset}
Our approach leverages multiple handwritten diagram datasets to ensure robustness across different diagram types and drawing styles. We employed a strategic dataset partitioning scheme where different subsets serve specific model training purposes.

\subsection{Dataset Sources}

We collected data from seven diverse sources, encompassing various diagram types commonly found in technical documentation and educational materials:

\begin{enumerate}
	\item \textbf{Finite Automata Dataset} \cite{online1}: Graph diagrams from the Czech Technical University collection, focusing exclusively on finite state machine representations.
	\item \textbf{Flowchart Dataset (FCB)} \cite{online3}: Flowchart diagrams from the Czech Technical University repository.
	\item \textbf{OHFCD\_1 Dataset} \cite{ohfcd}: Handwritten flowchart diagrams from the Computer Vision Center at UAB.
	\item \textbf{hdBPMN Dataset} \cite{BPMN}: Business Process Model and Notation diagrams from the Data and Web Science Lab.
	\item \textbf{Handwritten UML Class Diagrams} \cite{modelsketch}: Class diagram collection from Kaggle.
	\item \textbf{CircuitNet Dataset} \cite{circuitnet}: Electronic circuit diagram repository.
	\item \textbf{AI2D Dataset} \cite{ai2d}: Educational science diagrams from the Allen Institute for AI.
\end{enumerate}

\subsection{Dataset Utilization Strategy}

We implemented a targeted approach for dataset utilization based on the specific requirements of our two-stage architecture:

\textbf{Classifier Training:} To enhance model robustness and generalization capabilities, we employed class diagrams, circuit diagrams, and school diagrams from the UML, CircuitNet, and AI2D datasets respectively. This diverse selection ensures the classifier can effectively distinguish between different diagram types across various domains.

\textbf{Extractor Training:} For the specialized extraction tasks, we focused exclusively on flowchart and graph diagrams from the FCB, OHFCD\_1, and Finite Automata datasets. This targeted approach allows the extraction models to achieve high precision on the specific structural elements relevant to these diagram types.
